---
title: "Day 3 Homework"
output: html_document
---

The goal of this homework is to get practice with manipulating data.

#  My main questions are: 
  - how to enter new factor columns conditionally based on values in other columns
  - the use of the dcast function
  - how to remove all "<" and ">" characters from the entire data.frame
  

1. Make both a "wide" and a "long" version of your data

```{r}

install.packages("reshape2")
library(reshape2)

Mydata <- read.csv("UNStats.csv", na.strings = c("NA", "N/A", "n/a", " ", "..", "—"))
head(Mydata)
colnames(Mydata)
str(Mydata2)

#This is to remove the "<" and ">" but I can't get it to work. Is there another way?: 
gsub(">", "", Mydata)
gsub("<", "", Mydata)

#Changing the column names (could also do one long list, but easier to cut/paste this way:
names(Mydata)[names(Mydata)=="X"]<- "Country"
names(Mydata)[names(Mydata)=="Total.population..thousands."]<- "Population"
names(Mydata)[names(Mydata)=="Share.of.persons.aged.65.years.or.over..percentage."]<- "%_over_65"
names(Mydata)[names(Mydata)=="Under.five.mortality..deaths.per.1.000.live.births."]<- "Child_Mortality"
names(Mydata)[names(Mydata)=="Life.expectancy.at.birth..years."]<- "Life_Expectancy"
names(Mydata)[names(Mydata)=="Adolescent.fertility.rate..births.per.1.000.women."]<- "Fertility_Rate"
names(Mydata)[names(Mydata)=="Contracep..tive.prevalence..percent..age."]<- "Contraceptive_Rate"
names(Mydata)[names(Mydata)=="Urbanization..percentage.of.total.population."]<- "Urbanization"
names(Mydata)[names(Mydata)=="Gross.domestic.product.at.PPP..interna..tional.dollars.per.capita."]<- "GDP_Percapita"
names(Mydata)[names(Mydata)=="Poverty..below..1.25..PPP..a.day..percent..age."]<- "Poverty"
names(Mydata)[names(Mydata)=="Gini.index.b..percent..age."]<- "Gini"
names(Mydata)[names(Mydata)=="Mean.years.of.schooling..years."]<- "Years_Schooling"
names(Mydata)[names(Mydata)=="Energy.use..kilograms.of.oil.equivalent.per.capita."]<- "Energy_Use"
names(Mydata)[names(Mydata)=="Carbon.dioxide.emissions..metric.tons.per.capita."]<- "CO2_Emmissions"
names(Mydata)[names(Mydata)=="Sulfur.dioxide.emissions..kilograms.per.capita."]<- "SO2_Emmissions"
names(Mydata)[names(Mydata)=="Particulate.matter.concentra..tion..urban.population.weighted.PM10.micrograms.per.cubic.meter."]<- "PM10_Emmissions"
head(Mydata)

Widedata <- Mydata

Longdata <- melt(Mydata, id.vars = 1)

```

2. Transform a variable

```{r}

#original data was in thousands, so this converts to actual: 

Mydata$Population_Actual <- Mydata$Population * 1000
head(Mydata)

```

3. Recode a factor

```{r}
levels(Mydata$Country)
Mydata$Country <- factor(Mydata$Affluence, levels = c(1:nrow(Mydata)))
levels(Mydata$Country) 

```


4. Add a new factor, where the levels are conditional on the values of some other variable. E.g., a factor of "high", "medium", and "low" based on thresholds for another variable.

```{r}

Mydata$Affluence <- ifelse(as.numeric(Mydata$Poverty) > 5, as.factor("Poor"), as.factor("Rich"))
Mydata$Affluence <- factor(Mydata$Affluence, levels = c(1,2), labels = c("Poor", "Wealthy"))
head(Mydata)

#I'm having trouble with factors vs. characters vs. numeric

#This is how to do it

myris.melt$Dimension[myris.melt$measurement.type %]


```

5. Reorder the levels of this new factor

```{r}
#assuming these were factors it would be: 

Mydata$Affluence <- factor(Mydata$Affluence, levels = c("Wealthy", "Poor"))
head(Mydata)

```

6. Choose a factor that could be re-coded as multiple dimensions (recall the sepal/petal vs. length/width breakdown). If you don't have one, try to create one.  Now create two (or more) new factors that represent these dimensions.

```{r}
levels(Widedata$Country)
Widedata$Region[1:57] <- factor("One")
Widedata$Region[58:100] <- factor("Two")
Widedata$Region[101:199] <- factor("Three")

```


7. Using `dcast`, aggregate some variable into "cell means" of some kind.

```{r}
colnames(Longdata)
MydataCast <- dcast(Longdata, Country ~ variable)
colnames(Longdata)

```

# note on casting with a LOT of variables
```{r}
mydata.wide <- dcast(mydata.long, [80 columns here] ~ split) # instead of having to type all the colnames here, use the following sequence to create a formula to insert those names: 

columns.to.keep <- colnames(mydata.long)[1:80]
columns.to.keep.pasted <- paste(columns.to.keep, collapse = " + ") # normally you give paste seperate arguments to paste togeter. THe collapse function lets you create one string of col 1 + col 2 etc. THen: 
my.cast.formula <- as.formula(paste(columns.to.keep.pasted, " ~ split"))
class(my.cast.formula) # forumla you can use
mydata.wide <- dcast(mydata.long, formula = my.cast.formula)




```



8.  Save all of the above in a well-documented R Markdown document.  Email me the .Rmd file (and if you use different or updated data from Day 2 Homework, send that to me, too).
