---
title: "Day 6 Homework"
output: html_document
---

1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data.  

```{r}

mydata <- read.csv("UNStats.csv", header=TRUE, na.strings = c("NA","na",".."," .. ","..."," ... ","-"," - ","--"," -- ")) 
str(mydata)
mydata

```


2. Write out (in plain English, in the notebook) what you are trying to do

# I want to go through each column and remove any unusual characters. Eg. greater than and equal signs. 

3. Write some "pseudo-code" comments in an R chunk, as an outline

-select column number
-apply sub function
-add new vector back to data.table
-create final data.table

4. Write some (real) code that will do what you want ONCE, on one subset of the data.

```{r}
column <- mydata[["x"]]
mydatarow1 <- gsub("[[:punct:]]","", mydata[["X"]])

```

5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.

```{r}
my.answer <- vector(length = 199)
for(this.column in 1:16){
    this.vector <- mydata[,mydata$this.counter]
    my.answer[this.counter] <- gsub("[[:punct:]]","",(this.vector))
}

```

6. Copy that loop again, and at a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.

```{r}
my.answer <- vector(length = 199)
for(this.column in 1:16){
    this.vector <- ifelse(class(mydata$this.column) = c("int", "num"), mydata[,mydata$this.counter], NULL) 
    my.answer[this.counter] <- gsub("[[:punct:]]","",(this.vector))
}

# I know something is wrong with this, but can't figure out what. 

```

