---
title: "Day 9 Homework"
output: html_document
---

1. Write a function to do a bunch of clean-up and formatting to your data.

```{r}
library(dplyr)
library(reshape2)

mydata <- read.csv("UNStats.csv", header=TRUE, na.strings = c("NA","na",".."," .. ","..."," ... ","-"," - ","--"," -- "))

colnames(mydata) <- c("country", "population", "elderly", "young", "fertility", "contraceptive", "urbanization", "GDP", "poverty", "Gini", "schooling", "energy", "CO2", "SO2", "P10")

rmsymbol <- function(x) { 
  columns <- ncol(x)
  for(i in columns) {
      x$i <- gsub("[[:punct:]]","", x[["i"]])
  }
  rows <- nrow(x)
  for(r in rows){
  summarize(x, x$NAs <- anyNA(r,))
  }
}
    
```

2. Make sure this function works on your data.

3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?

```{r}

rmsymbol(chickwts)

# no there is something wrong with the nrows function, but can't figure it out.

```

4. Fix your function to be able to work on the new data.

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.

```{r}

# To start, I would like either to remove all NAs or group by those with and without NAs. I'm trying two ways: 
rows <- c(1:nrow(mydata))
summarize(mydata, mydata$NA = anyNA(rows)

# But because the above isn't working, I'll just use: 

clean <- na.omit(mydata)

#then I can add summary stats: 

clean
is.numeric(clean$energy)
mutate(clean, energy_2 = energy/2)

# still getting the missing value where TRUE/FALSE needed error I asked about in class. Still not sure why. WHen I use a sample data set with numerical column vectors it works fine. And my clean data set is also numerical columns with no NAs. 

mtcars
mutate(mtcars, displ_1 = disp/61.0237)

# so using the mtcars data set, here is a possible function to get a summary table with the mean of all rows in each column if values are not characters: 

mean.cols <- function(x){
    row1 <- colnames(x)
    columns <- ncol(x)
  for(c in columns){
    row2 <- c(1:columns)
    coltotal <- x$c
    if (class(coltotal) == !character)  {
    mean(x$c)
      } else { NULL }
    }
  output <- data.frame(row1, row2)
}

mean.cols(mtcars)

```


